JD notes on effects

### 2022-07-03 manuscript/body.tex
trying very hard ☺

To make an outcome plot, we use a \emph{focal} predictor on the x-axis. The resulting plot will also depend on choices we make about other (non-focal) predictor variables. In ordinary (Gaussian) linear regression, the non-focal choices may have a simple additive effect on the outcome estimate. However, when the focal predictor has interactions, or when there is a non-linear link function, non-focal choices can also affect the slope of the response. Additional challenges arise when dealing with generalized linear (mixed) models (GL(M)Ms) which extend ordinary linear models to incorporate random effects. How we treat (include or exclude) the random effects also affect the slope and nature of the response.

@@

\dya{Simple OLS regression models are flexible and can easily be generalized, as generalized linear (mixed) models (GL(M)Ms), to examine more complex relationships, including nonlinear relationships between response and input variables e.g., higher order interactions via polynomials and splines, nonlinear transformation on the input variables, and nonlinear transformations via link functions. This flexibility comes with additional complications, for example, complex multivariate models may risk misinterpretation, and miscalculation of quantities of interest \citep{berry_improving_2012, leeper2017interpreting}. Further, coefficient estimates of GLMs involving nonlinear link functions lack direct interpretation\citep{leeper2017interpreting}, meaning that interpretation of derived quantities from these estimates requires some understanding of the specified model.}

@@

\dya{An alternative is to explore \emph{predictions} and \emph{effects} \citep{fox2009effect, leeper2017package, lenth2018package}.  For example, as a way to plan, public health officials may want to know the effects of household \code{income}, \code{wealthindex}, etc, on the predicted probability that slum dwellers will have improved water services.}

\jd{We should discuss before continuing this part.}
\BC{Okay. Some thoughts below your \P}
The outcome plots described above are often loosely called either prediction plots or effects plots. We endeavor here to make a conceptual distinction. If our goal is to \emph{predict} what we learn about the outcome variable by measuring the focal predictor, \NEW{then we would focus on the central (outcome) estimate which is the same for both prediction and effects plots. It simply tells us what we expect the outcome to be for any given value of the focal input variable. We may also want to provide uncertainty associated with the central estimate. How we choose to describe the uncertainty can inform the choice we make between prediction and effects plots. For instance, we may want to capture the uncertainty in the whole model (these include uncertainty due to intercept, focal and non-focal predictors, and random effects) or we may be interested in showing the uncertainty associated with the focal predictor only. In the former case, we use the prediction plots which generally focus on the \emph{total} effect, and so will often be suited for a univariate model. On the other hand, in the latter case, we introduce effects plots which focus on the direct effect of the focal predictor -- after controlling for other non-focal input variables -- from a multivariate model, and so will be preferable for multivariate models. Since effects plots focus on focal-specific uncertainty, they are usually narrower than prediction plots.} \BC{Not sure whether to say that uncertainties are described by the CIs?}

@@

The most common way of choosing the representative values is taking unique levels or quantiles of the focal input variable if discrete or continuous, respectively.  The predictions are then calculated by holding the non-focal input variables at their reference point (values chosen for the non-focal input variables, e.g., means), while varying the focal input variable, with the goal that the estimates represent how the model responds to the changes in the focal input variable \citep{fox2009effect, hanmer2013behind}. This procedure generates -- \emph{predictor effects} \citep{fox2009effect}, \emph{marginal predictions} \citep{leeper2017package} or \emph{estimated marginal means} \citep{lenth2018package}. 
\BC{We need to decide on how to proceed from here. I think we can refer to this as the central estimate. Re-writing and we can delete older version if that's okay.}
\NEW{In this article, we refer to these quantities as the central (outcome) estimates for an ``average case'' in a population; and the associated confidence intervals will either be a prerequisite for prediction or effects plots. In the case of effects plots, we can also choose an anchor (value chosen for the focal input variable). The concept of effects plots has been presented in the literature -- Shi et al. \citep{shi_evidence_2017} used effects plots to compare the difference in sexual risk behaviors between circumcised and uncircumcised men.}
\dya{In this article, we refer to these quantities as predictions and effects since they should, for example, tell us what we would expect the response to be at a particular value or level of the focal input variable, for an ``average case''. As mentioned above, in our case, the key difference between predictions and effects is in the computation of the confidence intervals. The concept of effects has been presented in the literature -- Shi et al. \citep{shi_evidence_2017} used effect-styled plots to compare the difference in sexual risk behaviors between circumcised and uncircumcised men.}

@@

\item \textbf{Anchor:} The value chosen for the focal input variable when estimating effect confidence intervals. The anchor choice does not affect the main prediction estimates. Typically chosen as the center point of the predictor variables corresponding to the focal input variable but other sensible values can be chosen too. If we are trying to focus on effects, the anchor from which we calculate confidence intervals becomes important. This is not true for the more conventional prediction confidence intervals. We can choose any anchor, e.g., zero-anchored (the value of the anchor is $0$), but center point seems like a natural, stable choice, we call this center-anchored. In Fig~\ref{fig:justify_anchors}, we compare effects plots anchored at the minimum and center (mean) of the focal input variable.

@@

\caption{{\bf A comparison of zero- and center- anchored effects.} The center-anchored effects are based on the mean of the focal predictor while the zero-anchored effects are based on the minimum of the focal predictor. As we mentioned earlier, the choice of the anchor do not affect the main estimates (central trend lines). \MOVE{Should me move this downstream}} 

>>>>

To make an outcome plot, we use a \emph{focal} predictor on the x-axis. The resulting plot will also depend on choices we make about other (non-focal) predictor variables. In ordinary (Gaussian) linear regression, the non-focal choices may have a simple additive effect on the outcome estimate. However, when the focal predictor has interactions, or when there is a non-linear link function, non-focal choices can also affect the slope of the response. Additional challenges arise when dealing with generalized linear (mixed) models (GL(M)Ms) which extend ordinary linear models to incorporate random effects. How we treat random effects can also the estimated response.

@@

@@

\jd{We should discuss before continuing this part.}
\BC{Okay. Some thoughts below your \P}
The outcome plots described above are often loosely called either prediction plots or effects plots. We endeavor here to make a conceptual distinction. If our goal is to \emph{predict} what we learn about the outcome variable by measuring the focal predictor, \NEW{then we would focus on the central (outcome) estimate which is the same for both prediction and effects plots. \jd{This \P\ should be about distinguishing -- what do they mean? I feel we can make good progress after a terminology meeting.}. It simply tells us what we expect the outcome to be for any given value of the focal input variable. We may also want to provide uncertainty associated with the central estimate. How we choose to describe the uncertainty can inform the choice we make between prediction and effects plots. For instance, we may want to capture the uncertainty in the whole model (these include uncertainty due to intercept, focal and non-focal predictors, and random effects) or we may be interested in showing the uncertainty associated with the focal predictor only. In the former case, we use the prediction plots which generally focus on the \emph{total} effect, and so will often be suited for a univariate model. On the other hand, in the latter case, we introduce effects plots which focus on the direct effect of the focal predictor -- after controlling for other non-focal input variables -- from a multivariate model, and so will be preferable for multivariate models. Since effects plots focus on focal-specific uncertainty, they are usually narrower than prediction plots.} \BC{Not sure whether to say that uncertainties are described by the CIs?}

@@

The most common way of choosing the representative values is taking unique levels or quantiles \jd{I still don't understand where quantiles come in? What's an example of the use of quantiles?} of the focal input variable if discrete or continuous, respectively.  The predictions are then calculated by holding the non-focal input variables at their reference point (values chosen for the non-focal input variables, e.g., means), while varying the focal input variable, with the goal that the estimates represent how the model responds to the changes in the focal input variable \citep{fox2009effect, hanmer2013behind}. This procedure generates -- \emph{predictor effects} \citep{fox2009effect}, \emph{marginal predictions} \citep{leeper2017package} or \emph{estimated marginal means} \citep{lenth2018package}. 
\BC{We need to decide on how to proceed from here. I think we can refer to this as the central estimate. Re-writing and we can delete older version if that's okay.}
\jd{I had a suggestion in the notes.}
In this article, we refer to these quantities as the central (outcome) estimates for an ``average case'' in a population; and the associated confidence intervals will either be \jd{WHAT?? a prerequisite for} prediction or effects plots. In the case of effects plots, we can also choose an anchor (value chosen for the focal input variable). The concept of effects plots has been presented in the literature -- Shi et al. \citep{shi_evidence_2017} used effects plots to compare the difference in sexual risk behaviors between circumcised and uncircumcised men.
\jd{This is useful text, but also feels repetitive; why are we coming back to the “concept” of effects plots?}

@@

\item \textbf{Anchor:} The value chosen for the focal input variable when estimating effect confidence intervals. The anchor choice does not affect the main prediction estimates. Typically chosen as the center point of the predictor variables corresponding to the focal input variable but other sensible values can be chosen too. If we are trying to focus on effects, the anchor from which we calculate confidence intervals becomes important. This is not true for the more conventional prediction confidence intervals. We can choose any anchor, but the center point is a natural, stable choice, we call this.
\jd{Why do we need this in the definitions section?:} In Fig~\ref{fig:justify_anchors}, we compare effects plots anchored at the center (mean) of the focal input variable with \ldots.

@@

\caption{{\bf A comparison of zero- and center-anchored effects.} \jd{This explanation is not currently correct, but do we really need to explain that the zero anchor means that the anchor is at zero? Also, the real difference between anchors is for multi-parameter predictors? Do we have an example of that?} The center-anchored effects are based on the mean of the focal predictor while the zero-anchored effects are based on the minimum of the focal predictor. As we mentioned earlier, the choice of the anchor do not affect the main estimates (central trend lines). \MOVE{Should me move this downstream}} 
